Hashlattice: A distributed system for monotonic peer-to-peer synchronization

### 2022-11-20 paper/hl_paper.tex
fix spacing

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

@@

\begin{document}
  \maketitle

@@

    \newcommand{\partialto}{\hookrightarrow}

@@

    \newcommand{\checkelem}{\ensuremath{\mathsf{check\_elem}}}
    \newcommand{\join}{\ensuremath{\mathsf{join}}}

@@

      \newcommand{\hashlookup}{\ensuremath{\mathsf{hash\_lookup}}}
      \newcommand{\datacomputationlookup}{\ensuremath{\mathsf{data\_computation\_lookup}}}

@@

      \newcommand{\latticelookup}{\ensuremath{\mathsf{lattice\_lookup}}}
      \newcommand{\latticecomputationlookup}{\ensuremath{\mathsf{lattice\_computation\_lookup}}}

@@

      \newcommand{\hashput}{\ensuremath{\mathsf{hash\_put}}}

@@

      \newcommand{\evaldatacomputation}{\ensuremath{\mathsf{eval\_data\_computation}}}

@@

      \newcommand{\transport}{\ensuremath{\mathsf{transport}}}
      \newcommand{\evallatticecomputation}{\ensuremath{\mathsf{eval\_lattice\_computation}}}

@@

    Dependencies are defined as follows:
    \begin{itemize}
      \item A data computation depends on a data value if the corresponding \evaldatacomputation function calls \hashlookup for the data value's hash code.
      \item A data computation depends on another data computation if the corresponding \evaldatacomputation function calls \datacomputationlookup for the other data computation.
      \item A semilattice value depends on a data value if the corresponding \checkelem function calls \hashlookup for the data value's hash code.
      \item A semilattice value depends on a data computation if the corresponding \checkelem function calls \datacomputationlookup for the data computation.
      \item A semilattice value depends on another semilattice value if the corresponding \checkelem function calls \latticelookup for the other semilattice value's semilattice.
      \item A semilattice value depends on a semilattice computation if the corresponding \checkelem function calls \latticecomputationlookup for the semilattice computation.
      \item A semilattice computation depends on a data value if the corresponding \evallatticecomputation function calls \hashlookup for the data value's hash code.
      \item A semilattice computation depends on a data computation if the corresponding \evallatticecomputation function calls \datacomputationlookup for the data computation.
      \item A semilattice computation depends on a semilattice value if the corresponding \evallatticecomputation function calls \latticelookup for the semilattice value's semilattice.
      \item A semilattice computation depends on another semilattice computation if the corresponding \evallatticecomputation function calls \latticecomputationlookup for the other semilattice computation.
    \end{itemize}

@@

    For semilattice values and semilattice computations, the database stores not only the value, but also a directed acyclic graph Merkle tree of their dependencies on semilattices or semilattice computations. This enables synchronization between different databases, since it is possible to check if any dependencies differ before joining the values. Synchronization proceeds recursively; within the \join call that merges two semilattice values, the \latticelookup and \latticecomputationlookup functions recursively compute a merged result from the two databases, using the Merkle trees as a frozen pointer to the relevant subset of each database at the time of the merge.

>>>>

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xspace}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

@@

\begin{document}
  \newcommand{\checkelem}{\ensuremath{\mathsf{check\_elem}}}
  \newcommand{\join}{\ensuremath{\mathsf{join}}}
  \newcommand{\partialto}{\hookrightarrow}
  \newcommand{\hashlookup}{\ensuremath{\mathsf{hash\_lookup}}}
  \newcommand{\datacomputationlookup}{\ensuremath{\mathsf{data\_computation\_lookup}}}
  \newcommand{\latticelookup}{\ensuremath{\mathsf{lattice\_lookup}}}
  \newcommand{\latticecomputationlookup}{\ensuremath{\mathsf{lattice\_computation\_lookup}}}
  \newcommand{\hashput}{\ensuremath{\mathsf{hash\_put}}}
  \newcommand{\evaldatacomputation}{\ensuremath{\mathsf{eval\_data\_computation}}}
  \newcommand{\transport}{\ensuremath{\mathsf{transport}}}
  \newcommand{\evallatticecomputation}{\ensuremath{\mathsf{eval\_lattice\_computation}}}

  \maketitle

@@

@@

@@

@@

@@

@@

@@

@@

    Dependencies are defined as follows:
    \begin{itemize}
      \item A data computation depends on a data value if the corresponding \evaldatacomputation{} function calls \hashlookup{} for the data value's hash code.
      \item A data computation depends on another data computation if the corresponding \evaldatacomputation{} function calls \datacomputationlookup{} for the other data computation.
      \item A semilattice value depends on a data value if the corresponding \checkelem{} function calls \hashlookup{} for the data value's hash code.
      \item A semilattice value depends on a data computation if the corresponding \checkelem{} function calls \datacomputationlookup{} for the data computation.
      \item A semilattice value depends on another semilattice value if the corresponding \checkelem{} function calls \latticelookup{} for the other semilattice value's semilattice.
      \item A semilattice value depends on a semilattice computation if the corresponding \checkelem{} function calls \latticecomputationlookup{} for the semilattice computation.
      \item A semilattice computation depends on a data value if the corresponding \evallatticecomputation{} function calls \hashlookup{} for the data value's hash code.
      \item A semilattice computation depends on a data computation if the corresponding \evallatticecomputation{} function calls \datacomputationlookup{} for the data computation.
      \item A semilattice computation depends on a semilattice value if the corresponding \evallatticecomputation{} function calls \latticelookup{} for the semilattice value's semilattice.
      \item A semilattice computation depends on another semilattice computation if the corresponding \evallatticecomputation{} function calls \latticecomputationlookup{} for the other semilattice computation.
    \end{itemize}

@@

    For semilattice values and semilattice computations, the database stores not only the value, but also a directed acyclic graph Merkle tree of their dependencies on semilattices or semilattice computations. This enables synchronization between different databases, since it is possible to check if any dependencies differ before joining the values. Synchronization proceeds recursively; within the \join{} call that merges two semilattice values, the \latticelookup{} and \latticecomputationlookup{} functions recursively compute a merged result from the two databases, using the Merkle trees as a frozen pointer to the relevant subset of each database at the time of the merge.
